---
title: "diagExamples"
author: "Rachel Wilkerson"
date: "22 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
load(file = "exampleCHDS.Rdata")
library(rlang);library(tidyverse);library(ggplot2);library(partitions);library(DiagrammeR);library(randomcoloR)
```

## CHDS example

The CHDS data here is supplied by Rodriguo Collazo. The CEG.AHC is found in Lorna Barclay's thesis.
```{r, eval=FALSE}
chds.df <-read.csv(file = "CHDS.latentexample1.csv")# #note: sst requires factors
chds.sst <- CEG.AHC(chds.df)
```

The staging is found in the results. Results shows the incoming edges contributing to each of the stages.
```{r}
chds.sst$result
```

These functions take the output of Lorna's code and transform it to run in Rachel's functions. The stage.key is essentially a tidyverse version of the output from results, and struct is the data at each of the stages. 

```{r, eval=FALSE, warning=FALSE}
#translate to run the diagnostic code 
tostagekey(chds.df, chds.sst)-> chds.stage.key
pull(map_df(chds.stage.key, ~distinct(.x,stage)),stage )-> chds.stages#the stages
chds.struct <- to.struct(chds.df,chds.stage.key,chds.sst) #the struct. a weird name for the data results of the AHC alg
chds.cuts <- colnames(chds.df)

```

The code below renders the CEG. $w_3$ and $w_4$ are in the same stage, but not the same position.
```{r}
renderCEG(chds.stage.key,chds.df)
```


This reproduces CEG B as seen in Collazo, Gorgen, Smith book.  
with a Bayes factor score of -2478.49.

####Global monitors 

The global monitor from Lauritzen is simply the Bayes factor. 

We can see the usefulness of the global monitor by examining the BF of models with different ordering.
```{r}
chds.sst2 <- CEG.AHC(chds.df[,c(2,1,3,4)]) #varorder E S L H 
chds.sst2$lik; #equivalent to chds.sst so same likelihoods 
chds.sst2.sk <- tostagekey(chds.df[,c(2,1,3,4)],sst = chds.sst2)
renderCEG(chds.sst2.sk, chds.df[,c(2,1,3,4)])

```

```{r}
chds.sst3 <- CEG.AHC(chds.df[,c(1,3,2,4)])
chds.sst3$lik #same likelihoods because statistically equivalent
chds.sst3.sk <- tostagekey(chds.df[,c(1,3,2,4)],sst = chds.sst3) # var order S L E H 
renderCEG(chds.sst3.sk, chds.df[,c(1,3,2,4)])
```

```{r}
chds.sst4 <- CEG.AHC(chds.df[,c(3,1,2,4)])
chds.sst4$lik #same likelihood because statistically equivalent
chds.sst4.sk <- tostagekey(chds.df[,c(3,1,2,4)],sst = chds.sst4) # var order L S E H 
renderCEG(chds.sst4.sk, chds.df[,c(3,1,2,4)])
```

These examples reproduce the results from the CEG book. 

This next set of function shapes the data and finds the reference prior. Here we take the effective sample size as 3, the highest number of levels we have for $X_l$.

```{r}
map(chds.stage.key[-1], ~select(.x,-key))->chds.sk#a little bit of finagling 
chds.sk <- c(chds.stage.key[1],chds.sk)
map(chds.sk, ~select(.x,-color))->chds.sk.nocol
chds.prior <-get.ref.prior(df=chds.df,struct=chds.struct,cuts=chds.cuts,stage.key=chds.sk.nocol,stages=chds.stages)#check that we can get the prior
```

#### Component monitor

The component monitor is the contribution from each of the positions to the final Bayes Factor. Nodes with a particularly high contribution may indicate an ill fitting model. For instance, here we examine the contribution of the positions in cut 2, $X_e$ on the subsequent cut $X_l$. 
```{r}
as.data.frame(cbind(stages,components))
```


We see here that, aside from the root node, the edges outgoing from $w_1$ and $w_3$ represent a large contribution to the Bayes factor score. This makes sense, as that set of edges is the only one which contains two incoming pathways: patients who have $X_s$ = High and $X_e$ = (High, Low). 

####Partition monitor

The partition monitor is of the form found in Freeman paper. The plot computes $p(U_t | x^{t-1})$ throughout the data set. 

```{r,eval=FALSE}
chds.part.monitor <- part.monitor(rho = .7,epsilon = 1.2,df_cut = chds.df,which.cut = 3,stage.key = chds.sk.nocol,n.monitor = 860)
```

```{r}
chds.crrnt.stg.probs <- do.call("rbind", lapply(chds.part.monitor[[3]], "[[", 4)) #possible.coloring 3
chds.alt1.stg.probs <- do.call("rbind", lapply(chds.part.monitor[[3]], "[[", 1))#possible.coloring 10
chds.alt2.stg.probs <- do.call("rbind", lapply(chds.part.monitor[[3]], "[[", 2))#possible.coloring 11
chds.alt3.stg.probs <- do.call("rbind", lapply(chds.part.monitor[[3]], "[[", 3))#possible.coloring 13

#possible.colorings[[3]] #(HH, HL, LH) (LL)
#possible.colorings[[13]]#(HH) (HL, LH) (LL)

chds.part.df <- as.data.frame(cbind(5:860,chds.crrnt.stg.probs, chds.alt1.stg.probs, chds.alt2.stg.probs, chds.alt3.stg.probs))
colnames(chds.part.df) <- c("t", "Staging1", "Staging2", "Staging3", "Staging4")
chds.part.df %>% 
  gather(key, value, -t) %>%
  ggplot(aes(x=t, y=value, colour=key)) + geom_line()

```

The stages are all the partitions in the Hasse diagram that are one move away from the current staging: 

* Staging 1: (HH HL LH) (LL) 
* Staging 2: (HL HH) (LH) (LL) 
* Staging 3: (HH LH) (HL) (LL)
* Staging 4: (HL LH) (HH) (LL)

The staging given by (HH LH) (HL) (LL) is the preferred one, and the second most likely staging is (HH HL LH) (LL). 


#### Parent child monitors 

This checks individual contributions from one position on another. It is analgous to the prequential parent child monitor for the BN found in Cowell, Dawid, Lauritzen book 

Here we examine the impact of two parent stages, $w_3$ and $w_5$ on child $w_7$. The y axis shows the cumulative log penalty. 

```{r,eval=FALSE}
chds.w7w3.pach <- ceg.child.parent.monitor(df=chds.df,
                         target.stage = "w7",
                         target.cut = 4, 
                         condtnl.stage = "w3",
                         struct = chds.struct,
                         stage.key = chds.sk.nocol, 
                         stages=chds.stages,
                         prior=chds.prior,
                         n = 860,
                         learn = F)
plot(chds.w7w3.pach[,1])

chds.w7w5.pach <- ceg.child.parent.monitor(df=chds.df,
                                         target.stage = "w7",
                                         target.cut = 4, 
                                         condtnl.stage = "w5",
                                         struct = chds.struct,
                                         stage.key = chds.sk.nocol, 
                                         stages=chds.stages,
                                         prior=chds.prior,
                                         n = 860,
                                         learn = F)
```

```{r}
chds.pach <- as.data.frame(cbind(1:860,chds.w7w3.pach[,1],chds.w7w5.pach[,1]))
colnames(chds.pach) <- c('t','w7w3','w7w5')

chds.pach %>%
  gather(key,value, -t) %>% 
  ggplot(aes(x=t,y=value,colour=key))+geom_line()

```

This can be used to detect anomalies in the data: for instance, in places where the cumulative log penalty plateaus and then jumps up suddenly. 

### Points to consider

* The diangostics can be used to pinpoint the elements of the structure that most differ from a BN. In larger CEGs, this can be used to visualize an element of the network. 

* Monitoring the stage structure offers a quick visual check on the stability of the staging

* The parent child monitors can be used to identify which incoming contributions to the stage have the highest penalty. These may indicate places in which the graph needs additional refinement. 